---
title: "hw2-visual"
author: "Yiyi"
date: "2024-02-27"
output: html_document
---

```{r}
library(readxl)
hw1_data <- read_excel("D:/Multivariate_analysis/hw1-data.xlsx")
View(hw1_data)
data<-data.frame(hw1_data)
library(ggplot2)
library(GGally)
```

#Univariate visualization
#How is the GPA distributed in all the samples?
```{r}
ggplot(data, aes(x=GPA)) + geom_bar(fill="steelblue")
```
#The numerical distribution of GPA ranged from approximately 2.8 to 4.0, with GPA values having a maximum distribution between 3.7 and 3.8

#How is mental scores of gaming hours distributed in all the samples?
```{r}
ggplot(data, aes(x=Gaming_Hours_Per_Week)) + geom_bar(fill="steelblue")
```
#In the sample, the most students played in 5 and 8 hours per week, with 5 and 4 students, respectively.
#Only one student plays the most games for 20 hours a week. Only 2 students play only 2 hours a week.

#Bivariate visualization
#Is there a relationship between GPA and the level of English?
```{r}
ggplot(data, aes(x=English_Level, y=GPA)) + geom_point() + geom_smooth(method=lm)
```
#GPA values are concentrated between 3.00 and 4.00, while TOEFL scores are relatively more dispersed.
#Students with higher English level tend to have higher GPA,especially at the higher range of TOEFL scores.

#Is there a relationship between GPA and the Study hours per week?
```{r}
library(ggplot2)
ggplot(data, aes(x=Study_Hours_Per_Week,y=GPA)) + 
  geom_point() +
  labs(title="The Relationship Between GPA and Study Hours", x="Mental_Health", y="GPA")
```
#Almost all the students who studied for more than 10 hours got good grades of more than 3.5.
#Students whose study time is less than 10 hours will almost all have a GPA below 3.3 or even lower. It shows that learning time does have an significant correlation with GPA.

#Is there a relationship between GPA and the sport time?
```{r}
library(ggplot2)
ggplot(data, aes(x=Sport_Hours_Per_Week,y=GPA)) + 
  geom_point() +
  labs(title="The Relationship Between GPA and the sport time", x="Sport_Hours_Per_Week", y="GPA")
```
#There is no significant correlation between GPA and the sport time.

#Multivariate visualization

#scatterplot matrix
```{r}
library(car)
labs.data1 <- c("English_Level","Social_Integration","Study_Hours_Per_Week","Mental_Health","Social_Media_Hours_Per_Week","Gaming_Hours_Per_Week","Sport_Hours_Per_Week")
scatterplotMatrix(~English_Level+Social_Integration+Study_Hours_Per_Week+Mental_Health+Social_Media_Hours_Per_Week+Gaming_Hours_Per_Week+Sport_Hours_Per_Week | 
GPA, data=data, var.labels=labs.data1,cex.labels=0.7, data1="boxplot",smooth=FALSE,reg.line=FALSE,
pch=c(1,16),col=rep("steelblue",2), legend.plot=FALSE)
```

#What is the relationship between GPA and English Level, study hours and social integration?
```{r}
model <- lm(GPA ~ English_Level + Social_Integration + Study_Hours_Per_Week + Mental_Health + Social_Media_Hours_Per_Week + Gaming_Hours_Per_Week + Sport_Hours_Per_Week, data = data)
summary(model)
```

```{r}
library(rmarkdown)
library(knitr)
library(htmlwidgets)
render("D:/Multivariate_analysis/hw2_visual.Rmd", output_format = "html_document")
```

